---
title: "HW 03 - Great Lakes Fish Datasets"
author: "Kayode Ayegbusi"
date: "02/15/2023"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

```{r load-data, message = FALSE}
catch_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')

fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1


```{r catch-data-subset}
# select the columns of interest and save to overwrite the original dataframe object
catch_data <- catch_data %>% 
  select(year, lake, species, region, values)
```



```{r fish-stock-subset}
#select the columns of interest and save to overwrite the original dataframe object
fish_stock <- fish_stock %>%
  select(YEAR, MONTH, LAKE, STATE_PROV, SPECIES, WEIGHT)
```

### Exercise 2

```{r glimpse catch-data}
#review the selected columns stored in catch_data
glimpse(catch_data)
```

```{r glimpse fish-stock}
#review the selected columns stored in fish_stock
glimpse(fish_stock)
```


### Exercise 3

```{r filter-catch-data }
# I want to filter out other information in the catch_data of no interest and retain data of interest
catch_data <- catch_data %>%
  filter(region == "U.S. Total",
         species ==  "Chinook Salmon",
         lake == "Michigan",
         eval(TRUE))
catch_data
```


```{r filter-fish-stock}
# I want to filter out other information in the fish_stock of no interest and retain data of interest
fish_stock <- fish_stock %>%
  filter("MI" %in% c("MI","WI", "IL", "IN"),
        SPECIES == "CHS",
        LAKE == "MI",
        eval(TRUE))
fish_stock
```


### Exercise 4
There are `r nrow(catch_data)` observations and `r ncol(catch_data)` variables in the catch_data data set.
The first year in the catch_data is `r min(catch_data$year)`
The last year in the catch_data is `r max(catch_data$year)`

There are `r nrow(fish_stock)` observations and `r ncol(fish_stock)` variables in the fish_stock data set.
The first year in the catch_data is `r min(fish_stock$YEAR)`
The last year in the catch_data is `r max(fish_stock$YEAR)`

### Exercise 5


```{r stock-weight-yearly}
#turn the monthly weight of the fish stock to yearly weight using group_by function and summarise()
fish_stock_yearly <- fish_stock %>% 
  group_by(YEAR) %>% 
  summarise(STOCK_WEIGHT_YR = sum(WEIGHT, na.rm = T))

```

### Code chunks provided

Remember to change eval = FALSE, to eval = TRUE, when you are ready to run the code/knit the document.

```{r finalize-fish-stock, eval = TRUE}
fish_stock_yearly <- fish_stock_yearly %>% 
# rename the variable so that the two dataframes have thesame format that can be combined together
  rename(year = YEAR) %>% 
# use mutate to convert the weight from kg to lbs
  mutate(stock_weight_yr_lbs = as.integer(STOCK_WEIGHT_YR * 0.454 / 1000)) %>% # 1 lb = 1 kg * 0.454
  select(year, stock_weight_yr_lbs)
```

```{r finalize-catch-data, eval = TRUE}
catch_data <- catch_data %>% 
# rename the variable values to catch_weight_yr_lbs
rename(catch_weight_yr_lbs = values)

```

### Exercise 6


```{r left-join}
#use left_join function to join the two data sets together
left_join(catch_data, fish_stock_yearly)
```


The line of code produced a combined dataset that has seventeen rows (observations) and six columns (variables), this combination retains common observations in both datasets and eliminate observations in fish_stock_yearly not found in catch_data but retains observations in catch_data not found in fish_stock_yearly.
```{r right-join}
#use right_join function to join the two data sets together
right_join(catch_data, fish_stock_yearly)
```
The line of code produced a combined dataset that has fiftytwo rows (observations) and six columns (variables), this combination retains common observations in both datasets and eliminate observations in catch_data not found in fish_stock_yearly but retains observations in fish_stock_yearly not found in catch_data. 

```{r inner-join}
#use inner_join function to join the two data sets together
inner_join(catch_data, fish_stock_yearly)
inner_join(fish_stock_yearly, catch_data)
```

This line of code produced a combined dataset that only retains the observations in both datasets that matched and eliminates observation in catch_data that are not present in fish_stock_yearly and vice versa.

```{r full-join}
#use full_join function to join the two data sets together
full_join(catch_data, fish_stock_yearly)
```
This line of code produced a combination of all the observations in both datasets.

```{r anti-join-v1, eval = TRUE}
catch_data %>% 
  anti_join(fish_stock_yearly)
```

This code produced zero combination because every observation in catch_data are present in fish_stock_yearly data set

```{r anti-join-v2, eval = TRUE}
fish_stock_yearly %>% 
  anti_join(catch_data)
```
This code produced a dataset that retains only the observations that are present in fish_stock_yearly but absent in catch_data

### Exercise 7

```{r pivot-longer, eval = TRUE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "catch_weight_yr_lbs",
               cols = 5:6)
```

### Exercise 8 -.5

```{r final-visualization, eval = TRUE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "weight_lbs",
               cols = 5:6) %>% 
  ggplot(aes(x = year, y = weight_lbs, color = activity)) +
  geom_point() +
  geom_line() +
  theme_bw()
```


The trend reached it peak in the mid 1990s, specifically 1995 for catch_weight_yr_lbs dataset but have been decline ever since.
