---
title: "Lab 03 - Great Lakes Fish Stocking"
author: "Kayode Ayegbusi"
date: "21/02/2023"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

There are `r nrow(fish_stock)` observations  and `r ncol(fish_stock)` variables in fish_stock dataset.

Each row represent the amount of fish stocked on monthly basis with the corresponding year, state/region, lake, species,  age in month, and weight



### Exercise 2


Add code chunks as needed.
```{r eval = TRUE}
#create a new dataframe called salmon_stock using the filter function to filter three species of interest
salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("ATS", "CHS", "COS"),
#Only filters those with total fish_stock valeus reported
    !is.na(WEIGHT),
#Filters out fish stock from all lakes except from Ontario
     LAKE != "ON"
  )
```

```{r statistics-of-salmon-s6tock}
nrow(salmon_stock)
ncol(salmon_stock)
names(salmon_stock)
class(salmon_stock)
salmon_stock
```
There are 5131 number of rows and 8 variables in the Salmon stock.
The names of the variables and there respective data type are YEAR (numerical), MONTH (numerical), LAKE (Character), STATE_PROV (Character), SPECIES (Character), NO_STOCKED (numerical), AGEMONTH (numerical), WEIGHT(numerical)

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.
