---
title: "Lab 03 - Great Lakes Fish Stocking"
author: "Kayode Ayegbusi"
date: "21/02/2023"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

There are `r nrow(fish_stock)` observations  and `r ncol(fish_stock)` variables in fish_stock dataset.

Each row represent the amount of fish stocked on monthly basis with the corresponding year, state/region, lake, species,  age in month, and weight



### Exercise 2



```{r eval = TRUE}
#create a new dataframe called salmon_stock using the filter function to filter three species of interest
salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("ATS", "CHS", "COS"),
#Only filters those with total fish_stock values reported
    !is.na(WEIGHT),
#Filters out fish stock from all lakes except from Ontario
     LAKE != "ON"
  )
```

```{r statistics-of-salmon-s6tock}
nrow(salmon_stock)
ncol(salmon_stock)
names(salmon_stock)
class(salmon_stock)
```
There are 5131 number of rows and 8 variables in the Salmon stock.
The names of the variables and there respective data type are YEAR (numerical), MONTH (numerical), LAKE (Character), STATE_PROV (Character), SPECIES (Character), NO_STOCKED (numerical), AGEMONTH (numerical), WEIGHT(numerical)

### Exercise 3

```{r average-weight}
#calculate average weight of salmon stock and add it to the original data set
salmon_stock <- salmon_stock %>%
  mutate(
    AVG_WEIGHT = WEIGHT / NO_STOCKED * 1000
  )
```
The mutate() function is used to calculate the average weight of salmon stock and add the new created variable call AVG_WEIGHT to the original dataset by dividing the weight by number of stock per observation

### Exercise 4


```{r total-weight-and-salmon-per-event}
#calaculate mean stock of fish and mean weight of stock per event for each lake
salmon_stock %>% 
  group_by(LAKE) %>% 
  filter(YEAR > 2000) %>% 
  summarize(MEAN_NO_STOCKED = mean(NO_STOCKED, na.rm = TRUE), # mean() won't return an NA if any NAs are present
      MEAN_WEIGHT = mean(WEIGHT, na.rm = T)) %>% 
  # arrange in descending order to have the highest at the top
  arrange(desc(MEAN_NO_STOCKED))
```
Lake Huron has the largest mean fish per stocking event and mean weight per stocking event respectively
### Exercise 5

```{r create-new-data-set-for-Trout-species}
#create a new fish stock of Trout species only
fish_stock %>%
  filter(
    SPECIES %in% c("BKT", "BRN", "LAT", "RBT", "TRT"), 
#Only filters those with total fish_stock values reported
    !is.na(WEIGHT),
#Filters out fish stock from all lakes except from Ontario
     LAKE != "ON") %>% 
#create average fish weight using mutate
  mutate(AVG_WEIGHT = WEIGHT/NO_STOCKED) %>% 
#use summarise to calculate the  mean and total weight of Trout fish stock
  group_by(LAKE) %>% 
   summarize(MEAN_NO_TROUT = mean(NO_STOCKED, na.rm = TRUE), # mean() won't return an NA if any NAs are present
      MEAN_WEIGHT = mean(WEIGHT, na.rm = T)) %>% 
  # arrange in descending order to have the highest at the top
  arrange(desc(MEAN_NO_TROUT))
```
Lake Michigan has the largest mean fish per stocking event and mean weight per stocking event respectively
### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.
